A Handbook for Tenured Track Faculty in Computer Science in the US

### 2024-03-23 faculty.tex
updates

\subsection{Serving on Program Committees (PC)} This is a good way to get to know what is hot in your field, and to get to know people in your field.  It is also a good way to get your name out.
You should try to serve on the program committee of the top conferences in your field.

\textbf{Advice}: You can ask senior colleagues or even your PhD advisor/collaborators to recommend you to serve on a PC. You can also directly email the PC chairs and ask to serve on the PC.  For example, I directly email OOPSLA and PLDI PC chairs and asked to serve on the PC, and they accepted my request (I do have papers at OOPSLA and PLDI when I asked). I tried this with SE conferences  e.g., ICSE (and of course also nominate myself through the official form), and it never work, even though I have many papers at ICSE.

@@

In general, I enjoy serving on NSF panels much more than serving on PC. While the work is more intense and takes more time, I found it to be more rewarding and learned a lot from it. Since my first NSF panel in 2019, I have served 7 panels (at least once every year since '19) and I look forward to serving more.

@@

\subsection{Volunteer for more Leadership roles} Once you have served on a few PCs and NSF panels, you should aim for more leadership roles.  For example, you can volunteer to be a chair of certain sessions/tracks of a large conference, or even a general chair of a workshop or conference.  These roles are more time consuming, but they are also more rewarding (e.g., more people will know about you) and in general will help your career, especially your tenure case.

\section{Service to the Department}
Your department chair will assign you to some committees to help run the department.  Typically for tenure track faculty the services are lighter than tenured ones (e.g., less leadership roles). For example, I have served on the \emph{PhD and/or MS admission committee (adcom)}, \emph{faculty search committee}, and \emph{awards committee} when I was at UNL. Being a member of the PhD adcom is very common for new assistant professors, as it is a good way to recruit students and build your lab.

\section{Service to the University}
\section{Other Types of Services}
\paragraph{Writing Recommendation Letters} You will be asked to write a lot of recommendation letters, for both students and colleagues.

\textbf{Advice}: you should have a template for recommendation letters.  You can find many templates online.  You should also ask your colleagues to share their templates with you.  You should also ask your colleagues to share their recommendation letters with you.  You can learn a lot from reading other people's recommendation letters.

>>>>

\subsection{Serving on Program Committees (PC)} You should try to serve on the PC  of the \emph{top} conferences in your field. This is a good way to get to know what is hot in your field, and to get to know people in your field.  It is also a good way to get your name out.

\textbf{Advice}: You can ask senior colleagues or even your PhD advisor/collaborators to recommend you to serve on a PC. You can also directly email the PC chairs and ask to serve on the PC.

@@

@@

\subsection{Volunteer for more Leadership roles} Once you have served on a few PCs and NSF panels, you should aim for more leadership roles.  For example, you can volunteer to be a chair of certain sessions/tracks of a large conference, or even a general chair of a workshop or conference.  These roles are more time consuming, but they can be more rewarding (e.g., more people will know about you) and in general can help your career, especially your tenure case.

\subsection{Personal Experience} I have served on many PCs and NSF panels. The initial services are mostly through ``cold'' emails. For example, I directly email OOPSLA and PLDI PC chairs and asked, and they accepted my request (I did have papers at OOPSLA and PLDI when I asked). I tried this with SE conferences  e.g., ICSE (and also nominate myself through the official form), and it never works, even though I have many papers at ICSE.
I also emailed NSF program managers and asked to serve on a panel, and they accepted my request.

I felt that I didn't get much out of serving on PC and didn't really like it much. However, I still do it because I feel I should give back to the community.
On the otherhand, I throroughly enjoy serving on NSF panels. While the work is more intense and takes more time, I found it more rewarding and enjoy the discussions. Since my first NSF panel in 2019, I have served 7 panels (at least once every year since '19) and I look forward to serving more.

I have also volunteered for leadership roles, e.g., I directly asked the general chair of ICSE'21 to see if I can take more leadership roles and was then invited to serve as a proceeding co-chair. In hindsight this particular role is a lot of work and I didn't really like many aspects of it, but it was an eye-openning experience.

\section{Service to the Department}
Your department chair will assign you to some committees to help run the department.  Typically for tenure track faculty the services are lighter than tenured ones (e.g., less leadership roles). Common committees include PhD and/or MS admission, faculty search, and colloquium committees.

Depends on your preference and how you are perceived, you can take more leadership roles when you are close to tenure, such as being the chair of a committee or chair/director of a program.  These roles are more time consuming and might involve some politics, but they are also more rewarding and can help your tenure case.

\textbf{Advice}: You should try to be strategic about which committees you serve on that can help build your lab and tenure case. For example, serving in the PhD adcom is very common for new assistant professors, as it is a good way to recruit students for your lab. Serving on the faculty search committee is also a good way to recruit new faculty members that you like (and also let you know what the job market is like, just in canse you want to move). Serving on the colloquium committee is also a good way to invite speakers in your field that can be your \emph{tenure letter writers}.

\subsubsection{Personal} Since coming to GMU, I have been much more active in serving the department.  I share the vision of the department and want to help achieve it. In addition to the typical services load, I volunteer to serve in additional committees and leadership roles. For example, I maintain the CSRankings DB for the department, I help organize the Open House to recruit admitted students, I serve in the Dept Chair renewal committee, etc.
Even though I am not on the search committee, I speak to every candidate and help recruit strong ones. I agree to be director of the MS program, which is a lot of work and typically done by tenured faculty.

Sometimes I feel overwhelmed by the services, but I also feel that I am making a difference in the department and that is rewarding.
I am also building a good relationship with my colleagues and the department chair, and that is important for my tenure case.
In addition, I learn a lot from being in various roles (e.g., I learned how class scheduling work, how the deparmtnet is run, how to recruit students, how to recruit faculty, how to run a colloquium, etc) and feel more confident in my leadership skills.

In short, if you are enthusiastic about the department, want to make a difference, and \emph{feel comfortable at handling additional load}, you can volunteer for more services and leadership roles.

\section{Service to the University}
\section{Other Types of Services}
\paragraph{Writing Recommendation Letters} You will be asked to write recommendation letters for both students and colleagues.

\textbf{Advice}: %you should have a template for recommendation letters.  You can find many templates online.  You should also ask your colleagues to share their templates with you.  You should also ask your colleagues to share their recommendation letters with you.  You can learn a lot from reading other people's recommendation letters.
